EGGs are adhesive!

### 2025-03-04 calco/main.tex

The previous lemma allows us to deduce the following results in an $\mathcal{M}$-adhesive context.
\todo{A questo punto la proposizione seguente possiamo anche buttarla via se serve recuperare spazio}
\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\noindent
\parbox{7.5cm}{
	\begin{restatable}{proposition}{mps}\label{prop:kerset}
		Suppose that in $\Set$ the cube aside is given, whose top and bootom faces are pushout and such that $m\colon A\mto C$ and $m'\colon A'\mto C'4$ are injections. Then the right square, made by the kernel pairs of the vertical arrows is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\begin{example}\label[example]{lab_2}
	Consider again $\Sigma$ the signature of \Cref{exa_3}, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled by a morphism
	$f: \mathcal{G} \to \mathcal{G}^\Sigma$ that is characterised by the image of the edges. 
	If $f(h(1) = a$, $f(h_2) = 2$, $f(h_3) = *$, and $f(h_3) = /$, we represent it visually 
	by putting the identifiers of the edges in  $\mathcal{G}^\Sigma$ inside the boxes representing the edges of $\mathcal{G}$
	as
		\begin{center}
		\begin{tikzpicture}
			\begin{pgfonlayer}{nodelayer}
				\node[style=none] (lab1) at (-1.8, 0.8) {$h_1$};
				\node[style=small box] (h1) at (-1.8, 0.3) {$a$};
				\node[style=small box] (2) at (-1.8, -0.5) {$2$};
				\node[style=none] (lab2) at (-1.8, -1.0) {$h_2$};
				\node[style=none] (h1b) at (-1.8, 0.3) {};
				\node[style=none] (th2) at (0, 0.3) {};
				\node[style=none] (bh2) at (0, -0.3) {};
				\node[style=none] (h2b) at (0, 0) {};
				\node[style=none] (th3) at (1.8, 0){};
				\node[style=none] (bh3) at (1.8, -0.6){};
				\node[style=none] (h3c) at (1.8, -0.3){};
				\node[style=none] (2c) at (-1.8, -0.5){};
				\node[style=none] (lab3) at (0, 0.8) {$h_3$};
				\node[style=medium box] (h2) at (0, 0) {$*$};
				\node[style=none] (lab4) at (1.8, 0.5) {$h_4$};
				\node[style=medium box] (h3) at (1.8, -0.3) {$/$};
				\node[style=none](labv2) at (-0.9, -0.9) {$v_2$};
				\node[style=node] (v2) at (-0.9, -0.5) {};
				\node[style=none] (labv1) at (-0.9, 0.7) {$v_1$};
				\node[style=node] (v1) at (-0.9, 0.3) {};
				\node[style=none] (labv3) at (0.9, 0.4) {$v_3$};
				\node[style=node] (v3) at (0.9, 0) {};
				\node[style=none] (labv4) at (2.7, 0.1) {$v_4$};
				\node[style=node] (v4) at (2.7, -0.3) {};
			\end{pgfonlayer}
			\begin{pgfonlayer}{edgelayer}
				\draw (h1b.center) to (th2.center);
				\draw (2c.center) to (v2);
				\draw[in=180, out=30] (v2) to (bh2.center);
				\draw (h2b.center) to (th3.center);
				\draw (v4) to (h3c.center);
				%\draw (h3c) to (v4);
				\draw[in=180, out=-30](v2) to (bh3);
			\end{pgfonlayer}
        	\end{tikzpicture}
	\end{center}

	\iffalse
	\begin{center}\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_1$}] (A) at (0,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_2$}] (B) at (0,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_3$}] (C) at (3,-0.75) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_4$}] (D) at (3,-2.25) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=right:{$v_5$}] (E) at (6,-1.5) {};
			\draw[->-=.5] (1.75,-0.75)--(C)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (1.25, -1) rectangle (1.75, -0.5) {};
			\draw[->-=.5] (4.75,-1.5)--(E)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (-1.5,0)--(A)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (4.25, -1.75) rectangle (4.75, -1.25) {};
			\node at (4.5, -1.05){$h_3$};
			\node at (1.5, -0.3){$h_2$};
			\node at (-1.75, 0.45){$h_1$};
			\draw[rounded corners] (-2, -0.25) rectangle (-1.5, 0.25) {};
			\draw[->-=.5] (A)..controls(0.5,0)and(1.2,-0.2)..(1.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (B)..controls(0.5,-1.5)and(1.2,-1.3)..(1.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

@@

\begin{lemma}\label{lemma:stab}
	In $\EqHyp$, pushouts along regular monos are stable.
\end{lemma}

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, Q_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, Q'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and that suppose in the first diagram below,  all the vertical faces are pullbacks, the bottom face is a pushout and $h$ is a regular mono. By \Cref{prop:eqhyp_complete} the same is true for the other two cubes, by \Cref{cor:mono2}, $h_E$ and $h_V$ are monos and so the top faces of these cubes are pushouts.
	\[
	\xymatrix@C=10pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'} & &A_1'\ar[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]_{b_E} \ar[dl]_{z_E'} &&B_1'\ar[dd]|\hole_(.65){a_V}\ar@{>->}[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{z_V'} \\ 
		\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d} &&A_3'  \ar[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}
		&&B_3'  \ar[dd]_{c_V}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\
		&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{z} & &A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E} &&B_1\ar@{>->}[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{z_V} \\
		\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 && A_3 \ar@{>->}[rr]_{p_E} & & A_4&&B_3 \ar@{>->}[rr]_{p_V} & & B_4 }
	\]

	\noindent 
\parbox{10cm}{Now, if $d=(d_E, d_V, d_Q)$, we can pullback the third component to get the solid part of the cube aside. Notice, moreover that the commutativity of the solid diagram yields the existence of the dotted $w\colon T\to Y$. By the usual composition and decomposition properties of pullbacks (cfr.~\Cref{lem:pb1}) the left fact is a pullback. By the first point of \Cref{prop:eqhyp_complete} the bottom face is a pushout and $h_Q$ is a mono by \Cref{cor:mono2}, so the top face is a pushout too.}\hfill\parbox{3cm}{\xymatrix@C=10pt@R=10pt{&T\ar[dd]|\hole_(.65){x_2}\ar@{>->}[rr]^{x_1} \ar@{.>}[dl]_{w} && U \ar[dd]^{u_2} \ar[dl]_{u_1} \\ Y  \ar[dd]_{y_2}\ar@{>->}[rr]^(.7){y_1} & & Q_4' \ar[dd]_(.3){d_Q}\\&Q_1\ar@{>->}[rr]|\hole^(.65){h_C} \ar[dl]^{k_Q} && Q_2 \ar[dl]^{z_Q} \\Q_3 \ar@{>->}[rr]_{p_C} & & Q_4 }}

@@

@@

	Now, \Cref{lem:mpo}, let us take the kernel pair

@@

@@

@@

\noindent
\parbox{10cm}{
\begin{lemma}\label{lem:pb1}
	Let $\X$ be a category, and consider the diagram aside, in which the right square is a pullback. Then the whole rectangle is a pullback if and only if the left square is one.
\end{lemma}} 
	\parbox{4cm}{
	\xymatrix@R=18pt{X \ar[d]_{a} \ar[r]^{f}& \ar[r]^{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ A \ar[r]_{h}& B \ar[r]_{k}& C}}
\commentato{
\begin{proof}
	$(\Rightarrow)$ Let $q_1\colon Q\to Y$ and $q_2\colon Q\to A$ be two arrows such that $b\circ q_1=h\circ q_2$, if we compute we get
	\[c\circ g\circ q_1  =k\circ b\circ q_1=k\circ h\circ q_2\]
	\parbox{9cm}{Thus by the pullback property of the whole rectangle we get the dotted $l$ in the diagram on the side. 
	All we have to prove is that $f\circ l=q_1$. By construction we know that $g\circ f\circ l= g\circ q_1$, while we also have}
	\hfill \parbox{4cm}{\xymatrix@R=18pt{Q \ar@/_.3cm/[dr]_{q_2}\ar@/^.4cm/[rrr]^{g\circ q_1} \ar@{.>}[r]_{l} & X \ar[d]_{a} \ar[r]_{f}& \ar[r]_{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ & A \ar[r]_{h}& B \ar[r]_{k}& C}}
	\[	b\circ f\circ l = h\circ a\circ l= h\circ q_2=b\circ q_1\]
	and we can conclude since the right square in the original diagram is a pullback.

@@

\noindent
\parbox{7.5cm}{
\mps*}
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

>>>>

The previous result allows us to deduce the following lemma in an $\mathcal{M}$-adhesive context.

%\todo{A questo punto la proposizione seguente possiamo anche buttarla via se serve recuperare spazio}
\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

%\todo{'sta cragna che segue è falsa come una moneta da 3 euro: cancellare}
%\noindent
%\parbox{7.5cm}{
%	\begin{restatable}{proposition}{mps}\label{prop:kerset}
%		Suppose that in $\Set$ the cube aside is given, whose top and bottom faces are pushout and such that $m\colon A\mto C$, $m'\colon A'\mto C'$ and $g\colon C\mto D$ are injections. Then the right square, made by the kernel pairs of the vertical arrows, is a pushout.
%\end{restatable}}\hfill 
%\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar@{>->}[rr]_{g} & & D }}

@@

\begin{example}\label[example]{lab_2}
	Consider again $\Sigma$ the signature of \Cref{exa_3}, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled by a morphism
	$f: \mathcal{G} \to \mathcal{G}^\Sigma$ that is characterised by the image of the edges. 
	If $f(h(1) = a$, $f(h_2) = 2$, $f(h_3) = *$, and $f(h_3) = /$, we represent it visually 
	by putting the identifiers of the edges in  $\mathcal{G}^\Sigma$ inside the boxes representing the edges of $\mathcal{G}$
	as
		\begin{center}
		\begin{tikzpicture}
			\begin{pgfonlayer}{nodelayer}
				\node[style=none] (lab1) at (-1.8, 0.8) {$h_1$};
				\node[style=small box] (h1) at (-1.8, 0.3) {$a$};
				\node[style=small box] (2) at (-1.8, -0.5) {$2$};
				\node[style=none] (lab2) at (-1.8, -1.0) {$h_2$};
				\node[style=none] (h1b) at (-1.8, 0.3) {};
				\node[style=none] (th2) at (0, 0.3) {};
				\node[style=none] (bh2) at (0, -0.3) {};
				\node[style=none] (h2b) at (0, 0) {};
				\node[style=none] (th3) at (1.8, 0){};
				\node[style=none] (bh3) at (1.8, -0.6){};
				\node[style=none] (h3c) at (1.8, -0.3){};
				\node[style=none] (2c) at (-1.8, -0.5){};
				\node[style=none] (lab3) at (0, 0.8) {$h_3$};
				\node[style=medium box] (h2) at (0, 0) {$*$};
				\node[style=none] (lab4) at (1.8, 0.5) {$h_4$};
				\node[style=medium box] (h3) at (1.8, -0.3) {$/$};
				\node[style=none](labv2) at (-0.9, -0.9) {$v_2$};
				\node[style=node] (v2) at (-0.9, -0.5) {};
				\node[style=none] (labv1) at (-0.9, 0.7) {$v_1$};
				\node[style=node] (v1) at (-0.9, 0.3) {};
				\node[style=none] (labv3) at (0.9, 0.4) {$v_3$};
				\node[style=node] (v3) at (0.9, 0) {};
				\node[style=none] (labv4) at (2.7, 0.1) {$v_4$};
				\node[style=node] (v4) at (2.7, -0.3) {};
			\end{pgfonlayer}
			\begin{pgfonlayer}{edgelayer}
				\draw (h1b.center) to (th2.center);
				\draw (2c.center) to (v2);
				\draw[in=180, out=30] (v2) to (bh2.center);
				\draw (h2b.center) to (th3.center);
				\draw (v4) to (h3c.center);
				%\draw (h3c) to (v4);
				\draw[in=180, out=-30](v2) to (bh3);
			\end{pgfonlayer}
        	\end{tikzpicture}
	\end{center}
%
	\iffalse
	\begin{center}\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_1$}] (A) at (0,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_2$}] (B) at (0,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_3$}] (C) at (3,-0.75) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_4$}] (D) at (3,-2.25) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=right:{$v_5$}] (E) at (6,-1.5) {};
			\draw[->-=.5] (1.75,-0.75)--(C)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (1.25, -1) rectangle (1.75, -0.5) {};
			\draw[->-=.5] (4.75,-1.5)--(E)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (-1.5,0)--(A)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (4.25, -1.75) rectangle (4.75, -1.25) {};
			\node at (4.5, -1.05){$h_3$};
			\node at (1.5, -0.3){$h_2$};
			\node at (-1.75, 0.45){$h_1$};
			\draw[rounded corners] (-2, -0.25) rectangle (-1.5, 0.25) {};
			\draw[->-=.5] (A)..controls(0.5,0)and(1.2,-0.2)..(1.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (B)..controls(0.5,-1.5)and(1.2,-1.3)..(1.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

@@

\begin{lemma}\label{lemma:stab}
	In $\EqHyp$, $\reg(\EqHyp), \mathcal{N}$-pushouts are stable.
\end{lemma}

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, Q_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, Q'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and that suppose in the first diagram below,  all the vertical faces are pullbacks, the bottom face is a pushout $h$ is a regular mono and $k_Q\colon Q_1\mto Q_3$ is mono. By \Cref{prop:eqhyp_complete} the same is true for the other two cubes, by \Cref{cor:mono2}, $h_E$ and $h_V$ are monos and so the top faces of these cubes are pushouts.
	\[
	\xymatrix@C=10pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'} & &A_1'\ar[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]_{b_E} \ar[dl]_{z_E'} &&B_1'\ar[dd]|\hole_(.65){a_V}\ar@{>->}[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{z_V'} \\ 
		\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d} &&A_3'  \ar[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}
		&&B_3'  \ar[dd]_{c_V}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\
		&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{z} & &A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E} &&B_1\ar@{>->}[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{z_V} \\
		\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 && A_3 \ar@{>->}[rr]_{p_E} & & A_4&&B_3 \ar@{>->}[rr]_{p_V} & & B_4 }
	\]

	\noindent 
\parbox{10cm}{Now, if $d=(d_E, d_V, d_Q)$, we can pull back the third component to get the solid part of the cube aside. Notice, moreover that the commutativity of the solid diagram yields the existence of the dotted $w\colon T\to Y$. By the usual composition and decomposition properties of pullbacks (cfr.~\Cref{lem:pb1}) the left face is a pullback. By the first point of \Cref{prop:eqhyp_complete} the bottom face is a pushout and $h_Q$ is a mono by \Cref{cor:mono2}, so the top face is a pushout too.}\hfill\parbox{3cm}{\xymatrix@C=10pt@R=10pt{&T\ar[dd]|\hole_(.65){x_2}\ar@{>->}[rr]^{x_1} \ar@{>.>}[dl]_{w} && U \ar[dd]^{u_2} \ar[dl]_{u_1} \\ Y  \ar[dd]_{y_2}\ar@{>->}[rr]^(.7){y_1} & & Q_4' \ar[dd]_(.3){d_Q}\\&Q_1\ar@{>->}[rr]|\hole^(.65){h_C} \ar@{>->}[dl]^{k_Q} && Q_2 \ar[dl]^{z_Q} \\Q_3 \ar@{>->}[rr]_{p_C} & & Q_4 }}

@@

Now, $d_V$ is a regular epi, thus

@@

@@

	\begin{example}
		Un controesempio:
		Consideriamo i grafi senza edge e con nodi dati dal seguente cubo in cui le facce verticali sono pb e le due orizzontali sono po (tutte le frecce sono inclusioni)
		\[
	\xymatrix@C=10pt@R=10pt{&\emptyset \ar[dd]|\hole_(.65){\id{\emptyset}}\ar@{>->}[rr]^{?_2} \ar[dl]_{?_1} && 2 \ar[dd]^{\id{2}} \ar[dl]_{i_{2,3}} \\ 1  \ar[dd]_{i_{1,2}}\ar@{>->}[rr]^(.7){i_{1,3}} & & 3 \ar[dd]_(.3){i_{3,4}}\\&\emptyset \ar@{>->}[rr]|\hole^(.65){?_2} \ar@{>->}[dl]^{?_2} && 2 \ar[dl]^{j} \\2 \ar@{>->}[rr]_{j'} & & 4 }
		\]

		Qui abbiamo $j'(0)=2, j'(1)=3$ e  $i_{1,3}(0)=2$
		Come quozienti consideriamo il seguente cubo

		\[
		\xymatrix@C=10pt@R=10pt{&\emptyset \ar[dd]|\hole_(.65){\id{\emptyset}}\ar@{>->}[rr]^{?_1} \ar[dl]_{?_1} && 1 \ar[dd]^{\id{1}} \ar[dl]_{i_{1,3}} \\ 1  \ar[dd]_{i_{1,2}}\ar@{>->}[rr]^(.7){i_{1,3}} & & 3 \ar[dd]_(.3){\id{3}}\\&\emptyset \ar@{>->}[rr]|\hole^(.65){?_2} \ar@{>->}[dl]^{?_1} && 1 \ar[dl]^{k} \\2 \ar@{>->}[rr]_{k'} & & 3 }
		\]

@@

\begin{example}
	Considet the following hypergaphs with equivalence:
	\[\mathcal{G}:=(\emptyset, 2, 2, ?_{2}, ?_2, \id{2}) \quad \mathcal{H}:=(\emptyset, 1,1, ?_{1}, ?_1, \id{1}) \quad \mathcal{G}'=(\emptyset, 2, 1, ?_{2}, ?_{2}, !_{2}) \]

There is a morphism $d:=(\id{\emptyset}, !_2, \id{1})\colon \mathcal{G}'\to \mathcal{H}$. We can then construct the cube aside, where the bottom face is a pushout and the vertical ones pullbacks.
	\[
\xymatrix@C=30pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'}  \\ 
	\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}' \ar[dd]^(.3){d} \\
	&\mathcal{G}\ar[rr]|\hole^(.3){({\emptyset, !_{2}, !_2})} \ar[dl]_{\id{\mathcal{G}}} && \mathcal{H} \ar[dl]^{\id{\mathcal{H}}}  \\
	\mathcal{G} \ar[rr]_{({\emptyset, !_{2}, !_2})} & & \mathcal{H}  }
\]\todo{sposta di lato}

\todo{finisci}

@@

\noindent
\parbox{10cm}{
\begin{lemma}\label{lem:pb1}
	Let $\X$ be a category, and consider the diagram aside, in which the right square is a pullback. Then the whole rectangle is a pullback if and only if the left square is one.
\end{lemma}} 
	\parbox{4cm}{
	\xymatrix@R=15pt{X \ar[d]_{a} \ar[r]_{f}& \ar[r]_{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ A \ar[r]^{h}& B \ar[r]^{k}& C}}
\commentato{
\begin{proof}
	$(\Rightarrow)$ Let $q_1\colon Q\to Y$ and $q_2\colon Q\to A$ be two arrows such that $b\circ q_1=h\circ q_2$, if we compute we get
	\[c\circ g\circ q_1  =k\circ b\circ q_1=k\circ h\circ q_2\]
	\parbox{9cm}{Thus by the pullback property of the whole rectangle we get the dotted $l$ in the diagram on the side. 
	All we have to prove is that $f\circ l=q_1$. By construction we know that $g\circ f\circ l= g\circ q_1$, while we also have}
	\hfill \parbox{4cm}{\xymatrix@R=18pt{Q \ar@/_.3cm/[dr]_{q_2}\ar@/^.4cm/[rrr]^{g\circ q_1} \ar@{.>}[r]_{l} & X \ar[d]_{a} \ar[r]_{f}& \ar[r]_{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ & A \ar[r]_{h}& B \ar[r]_{k}& C}}
	\[	b\circ f\circ l = h\circ a\circ l= h\circ q_2=b\circ q_1\]
	and we can conclude since the right square in the original diagram is a pullback.

@@

The category of $\Set$ enjoys two other remarkable and useful properties; it is \emph{distributive} and \emph{extensive}. Distributivity amount to the following property: for every family $\{X_i\}_{i}$ and an object $Y$, the unique morphisms $\phi$ and $\psi $ fitting in the diagrams below, where $j_i$, $k_i$ and $h_i$ are coprojections, are isomorphisms.
\[\xymatrix@C=15pt@R=15pt{&Y\times X_i \ar@{>->}[dl]_{j_i} \ar@{>->}[dr]^{\id{Y}\times h_i} &&&X_i \times Y \ar@{>->}[dl]_{k_i} \ar@{>->}[dr]^{ h_i \times \id{Y}} \\ \Sum_{i\in I} Y\times X_i \ar@{.>}[rr]_{\phi}&& Y\times (\Sum_{i\in I}X_i) & \Sum_{i\in I}  X_i \times Y \ar@{.>}[rr]_{\psi}&&  (\Sum_{i\in I}X_i) \times Y }\]

\noindent 
\parbox{11.5cm}{\hspace{15pt} Extensivity means that given a family of objects $\{X_{i}\}_{i\in I}$ and a family of squares as the one on the right, where $j_i$ is a  coprojection, then all the squares are pullbacks if and only if $(Z, \{k_i\}_{i\in I})$ is a coproduct.}\hfill 
\parbox{2cm}{\xymatrix@R=15pt{Z_i \ar[r]_{k_i} \ar[d]_{f_i} & Z\ar[d]^{f} \\ X_i \ar@{>->}[r]^{j_i} & X}}

These two property together yields the following fact.

\begin{lemma}\label{lem:nodim}
	Let $f,g\colon A\rightrightarrows B+C$ be two arrows with a coproduct as codomain. Then $(A, \{j_i\}_{i=0}^
	3)$ is a coproduct, where $j_i\colon A_i\mto A $ is defined by the following pullback squares.
	\[\xymatrix@C=30pt{A_0 \ar@{>->}[r]^{j_0}  \ar[d]_{p_0}& A   \ar[d]^{(f,g)}& A_1 \ar@{>->}[r]^{j_1}  \ar[d]_{p_1} & A \ar[d]^{(f,g)} \\
		B\times B \ar@{>->}[r]_-{i_B\times i_B} & (B+C)\times (B+C) & B\times C \ar@{>->}[r]_-{i_B\times i_C} & (B+C)\times (B+C)}\]
	\[ \xymatrix@C=30pt{A_2 \ar@{>->}[r]^{j_2}  \ar[d]_{p_2}& A   \ar[d]^{(f,g)}& A_3 \ar@{>->}[r]^{j_3}  \ar[d]_{p_3} & A \ar[d]^{(f,g)} \\
		C\times B \ar@{>->}[r]_-{i_C\times i_B} & (B+C)\times (B+C) & C\times C \ar@{>->}[r]_-{i_C\times i_C} & (B+C)\times (B+C)}\]
\end{lemma}

\commentato{
\noindent
\parbox{7.5cm}{
\mps*}
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar@{>->}[rr]_{g} & & D }}

\begin{proof}
		Let $E'\subseteq C'$ be the complement of the image of $m'$ and $\iota\colon E'\mto B'$ be the corresponding inclusion. By \Cref{lem:po_set} we know that $(D', \{n', g'\circ \iota\})$ is a coproduct. Thus by \Cref{lem:nodim} we also know that there exists $K_0$, $K_1$, $K_2$ and $K_3$ fitting in the following four pullback diagrams and such that $\{K_d, \{j_i\}_{i=0}^3\}$ is a coproduct.
	\[\xymatrix@C=45pt{K_0 \ar@{>->}[r]^{j_0}  \ar[d]_{(p_0, q_0)}& K_d   \ar[d]^{(\pi^1_d, \pi^2_d)}& K_1 \ar@{>->}[r]^{j_1}  \ar[d]_{(p_1, q_1)} & K_d \ar[d]^{(\pi^1_d, \pi^2_d)} \\
		B'\times B' \ar@{>->}[r]_-{n'\times n'} & D'\times D' & B'\times E' \ar@{>->}[r]_-{n'\times (g'\circ \iota)} & D'\times D'}\]
	\[ \xymatrix@C=45pt{K_2 \ar@{>->}[r]^{j_2}  \ar[d]_{(p_2, q_2)}& K_d   \ar[d]^{(\pi^1_d, \pi^2_d)}& K_3 \ar@{>->}[r]^{j_3}  \ar[d]_{(p_3, q_3)} & K_d \ar[d]^{(\pi^1_d, \pi^2_d)} \\
		E'\times B' \ar@{>->}[r]_-{(g'\circ \iota) \times n'} & D'\times D' & E'\times E' \ar@{>->}[r]_-{(g'\circ \iota)\times (g'\circ \iota)} & D'\times D'}\]
